Embedding Intuitionistic into Classical Logic

### 2023-05-10 LPAR/LPAR-EIICL.tex
small adjustments

%used for d
\usepackage{todonotes}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\newcommand{\red}[1]{{\color{red} #1}}

@@

\begin{definition}
	A \emph{Kripke structure}  \sout{$\mathcal K = (W, (v_w)_{w\in W})$} \red{$\mathcal K = (\mathbf W, (v_w)_{w\in W})$} consists of a partially ordered set $\mathbf W = (W, \leq)$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that \sout{$v_u(A)\leq v_w(A)$} \red{$v_u(A) = 1$ entails $v_w(A) = 1$} for all worlds $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
	\end{itemize}
	We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulae with \emph{IPC} (Intuitionistic Propositional Calculus).
\end{definition}
There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.

@@

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set \sout{$W$} \red{$(W, \leq)$}, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
	\begin{itemize}
		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/x]\models\varphi$.
	\end{itemize}
	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$. $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u\in W$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
	We denote the set of valid formulae with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to propositional tautologies there are now quantified formulae which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.

@@

\begin{theorem}
	\label{thm:reduction-first-order-short}
	There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and \sout{$\varphi$}\red{$\varphi^\#$} contains a new binary predicate $E$ as well as a number of new function symbols. Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
\end{theorem}

@@

A proof can be found in Appendix~\ref{proof:prop-simplification}.

\begin{remark}
	It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic. Our contribution is giving an explicit bound to height and degree that depends only on a simple property of the considered formula.
\end{remark}

@@

\begin{theorem}\label{thm:prop-translation}
	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
	with flat $\mathcal R$ and implicational $\mathcal S$. Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$. For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional \sout{Variable}\red{variable} $A^\lambda$. Define $\mathcal R^\#$ as the union over all sets
	$$
		\left\{\bigwedge_iA_i^\lambda\to\bigvee_jB_j^\lambda\:|\:\lambda\in\Lambda\right\}
	$$
	for $\bigwedge_iA_i\to\bigvee_jB_j\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
	$$
		\left\{\left(A^{\lambda\psi} \to B^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $\mathcal P^\#$ as the union over all sets
	$$
		\left\{A^\lambda\to B^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	 for each propositional variable $A$ occurring in $\varphi$, encoding persistency. Then
	$$
		\varphi^\# := \left(\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge \mathcal P^\#\right)\to P^\epsilon
	$$
	is classically valid if and only if $\varphi$ is intuitionistically valid.
\end{theorem}
%\begin{theorem}
%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
%\end{theorem}

@@

@@

It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}, but we make it explicit. In particular, our translation enables a direct translation of counter-models. For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
The translation from the last section will serve as \red{a} starting point for this section.

@@

\begin{definition}\label{def:fo-translation}
	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
	\begin{itemize}
		\item $\bot^b = \bot$.
		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
		\item $(\varphi\circ\psi)^b = \varphi^b\circ\sout{\psi^u}\red{\psi^b}$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is a new bound variable.
		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
		\item $(\forall x\varphi)^b = \forall w(\sout{i}\red{b}\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
	\begin{align*}
		K(\varphi) := \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
	\end{align*}
	with e.g.:
	\begin{align*}
		\text{PartialOrder}(\preceq) := &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{DomainSubset}(u, w) := &\:\forall x(E(x, u)\to E(x, w))\\
		\text{World}(u) := &\:\exists xE(x, u)\\
		\text{DomainClosed}(u) := &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
		\text{Persistent}(u, w) := &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
	\end{align*}
	Then define
	\[\varphi^C:= K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b\] where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
\end{definition}

@@

Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulae $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
where $\vec r, s$ are new constant symbols, each formula in $\mathcal R^{bS}$ is of the form
$$\forall w \forall \vec x(s\preceq w\to\vec E(x, w)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)),$$
each formula $\psi\in\mathcal X^{bS}$ is of the form
$$
	\resizebox{\hsize}{!}{$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to (w\preceq f_\psi(\vec x, w)\to A(\vec a, f_\psi(\vec x, w))\to B(\vec b, f_\psi(\vec x, w)))\to C(\vec c, w))$}
$$
and each formula $\psi\in\mathcal Q^{bS}$ is of the form
$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec{\red{b}}, w))$$
with $f_\psi$ being a new function symbol.

@@

Even if we \sout{don't}\red{don't} have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
$$
	\bigwedge\{\forall u\forall x\forall\vec y(E(x, u)\to E(x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\}
$$
and Persistent$(u, w)$ with
$$
	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a predicate in $\varphi$}\}.
$$Putting all of this together yields the following translation:
\begin{theorem}\label{thm:fo-translation}
	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above. For each $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name. For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$. Consider a new binary predicate $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal R^\#$ by including
	$$
		\forall u\forall \vec x\left(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)\right)
	$$
 	for each flat clause
 	$
 		\forall \vec x\left(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j)\right)\in\mathcal R,
 	$
 	$\mathcal X^\#$ by including
 	$$
 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec x, u))\to B(\vec b, f_\psi(\vec x, u))\to C(\vec c, u)\right)
 	$$
 	for each implication clause
	$
		\psi = \forall \vec x\left(\left(A(\vec a)\to B(\vec b)\right)\to C(\vec c)\right)\in\mathcal X,
	$
	and $\mathcal Q^\#$ by including
	$$
		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec x, u))\right)\to B(\vec b, u)\right)
	$$
	for each implication clause
	$
		\psi = \forall \vec x\left(\forall y\left(A(\vec a)\right)\to B(\vec b)\right)\in\mathcal X.
	$
	Then $\varphi$ is intuitionistically valid if and only if
	$$
		\varphi^\# := K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
	$$
	is classically valid, where $s, \vec r$ are new constants introduced by Herbrandization of free variables,
	\begin{align*}
		K(\varphi) := &\:\bigwedge\left\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\right\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a predicate in $\varphi$}\right\}$.}
	\end{align*}
	The size of the obtained formula is linear in the input. However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary predicate has been extended to a $n+1$-ary predicate.
\end{theorem}

@@

We have benchmarked our first-order translation on version 1.1.2 of the ILTP problem set~\cite{iltp} by performing our translation and then examining the generated classical problem using Vampire~\cite{kovacs2013first}.
For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} were used. All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem. The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of \red{the} instance.

@@

In general, our approach falls short of the state of the art. 
However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$\sout{resp. $|\mathcal X| \geq 6$}.

@@

Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers. Notably, while there is big overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.

We conclude that our approach already has merits when $|\mathcal Q| = 0$ and $|\mathcal X|$ is small.
We see that encoding idempotence does help to solve some additional benchmarks, but in general does not improve performance.
However, idempotence is a quite powerful concept from a model-theoretic perspective. We suspect that a more clever encoding of this property could improve results. There are other restrictions on models we could consider, based on structural properties of the formula, e.g. the relation between occurrences of different atoms.

>>>>

@@

\begin{definition}
	A \emph{Kripke structure}  $\mathcal K = (\mathbf W, (v_w)_{w\in W})$ consists of a partially ordered set $\mathbf W = (W, \leq)$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that $v_u(A) = 1$ entails $v_w(A) = 1$ for all worlds $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
	\end{itemize}
	We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulae with \emph{IPC} (Intuitionistic Propositional Calculus).
\end{definition}
There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.

@@

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $(W, \leq)$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
	\begin{itemize}
		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/x]\models\varphi$.
	\end{itemize}
	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$. $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u\in W$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
	We denote the set of valid formulae with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to propositional tautologies there are now quantified formulae which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.

@@

\begin{theorem}
	\label{thm:reduction-first-order-short}
	There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and $\varphi^\#$ contains a new binary predicate $E$ as well as a number of new function symbols. Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
\end{theorem}

@@

A proof can be found in Appendix~\ref{proof:prop-simplification}.
\begin{remark}
	It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic. Our contribution is giving an explicit bound to height and degree that depends only on a simple property of the considered formula.
\end{remark}

@@

\begin{theorem}\label{thm:prop-translation}
	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
	with flat $\mathcal R$ and implicational $\mathcal S$. Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$. For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional variable $A^\lambda$. Define $\mathcal R^\#$ as the union over all sets
	$$
		\left\{\bigwedge_iA_i^\lambda\to\bigvee_jB_j^\lambda\:|\:\lambda\in\Lambda\right\}
	$$
	for $\bigwedge_iA_i\to\bigvee_jB_j\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
	$$
		\left\{\left(A^{\lambda\psi} \to B^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $\mathcal P^\#$ as the union over all sets
	$$
		\left\{A^\lambda\to B^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	 for each propositional variable $A$ occurring in $\varphi$, encoding persistency. Then
	$$
		\varphi^\# := \left(\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge \mathcal P^\#\right)\to P^\epsilon
	$$
	is classically valid if and only if $\varphi$ is intuitionistically valid.
\end{theorem}
%\begin{theorem}
%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
%\end{theorem}

@@

Note that it is straightforward to obtain an intuitionistic counter-model to $\varphi$ with frame $\Lambda$ from a classical counter-model to $\varphi^\#$.

@@

It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}, but we make it explicit. In particular, our translation enables a direct translation of counter-models. For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
The translation from the last section will serve as a starting point for this section.

@@

\begin{definition}\label{def:fo-translation}
	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
	\begin{itemize}
		\item $\bot^b = \bot$.
		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^b$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is a new bound variable.
		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
		\item $(\forall x\varphi)^b = \forall w(b\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
	\begin{align*}
		K(\varphi) := \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
	\end{align*}
	with e.g.:
	\begin{align*}
		\text{PartialOrder}(\preceq) := &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{DomainSubset}(u, w) := &\:\forall x(E(x, u)\to E(x, w))\\
		\text{World}(u) := &\:\exists xE(x, u)\\
		\text{DomainClosed}(u) := &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
		\text{Persistent}(u, w) := &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
	\end{align*}
	Then define
	\[\varphi^C:= K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b\] where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
\end{definition}

@@

Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulae $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
where $\vec r, s$ are new constant symbols, each formula in $\mathcal R^{bS}$ is of the form
$$\forall w \forall \vec x(s\preceq w\to\vec E(x, w)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)),$$
each formula $\psi\in\mathcal X^{bS}$ is of the form
$$
	\resizebox{\hsize}{!}{$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to (w\preceq f_\psi(\vec x, w)\to A(\vec a, f_\psi(\vec x, w))\to B(\vec b, f_\psi(\vec x, w)))\to C(\vec c, w))$}
$$
and each formula $\psi\in\mathcal Q^{bS}$ is of the form
$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec{b}, w))$$
with $f_\psi$ being a new function symbol.

@@

Even if we do not have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
$$
	\bigwedge\{\forall u\forall x\forall\vec y(E(x, u)\to E(x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\}
$$
and Persistent$(u, w)$ with
$$
	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a predicate in $\varphi$}\}.
$$Putting all of this together yields the following translation:
\begin{theorem}\label{thm:fo-translation}
	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above. For each $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name. For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$. Consider a new binary predicate $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal R^\#$ by including
	$$
		\forall u\forall \vec x\left(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)\right)
	$$
 	for each flat clause
 	$
 		\forall \vec x\left(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j)\right)\in\mathcal R,
 	$
 	$\mathcal X^\#$ by including
 	$$
 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec x, u))\to B(\vec b, f_\psi(\vec x, u))\to C(\vec c, u)\right)
 	$$
 	for each implication clause
	$
		\psi = \forall \vec x\left(\left(A(\vec a)\to B(\vec b)\right)\to C(\vec c)\right)\in\mathcal X,
	$
	and $\mathcal Q^\#$ by including
	$$
		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec x, u))\right)\to B(\vec b, u)\right)
	$$
	for each implication clause
	$
		\psi = \forall \vec x\left(\forall y\left(A(\vec a)\right)\to B(\vec b)\right)\in\mathcal X.
	$
	Then $\varphi$ is intuitionistically valid if and only if
	$$
		\varphi^\# := K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
	$$
	is classically valid, where $s, \vec r$ are new constants introduced by Herbrandization of free variables,
	\begin{align*}
		K(\varphi) := &\:\bigwedge\left\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\right\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a predicate in $\varphi$}\right\}$.}
	\end{align*}
	The size of the obtained formula is linear in the input. However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary predicate has been extended to a $n+1$-ary predicate.
\end{theorem}

@@

We have benchmarked our first-order translation on version 1.1.2 of the ILTP problem set~\cite{iltp} by performing our translation and then examining the generated classical problem using Vampire~\cite{kovacs2013first}.
For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} were used. All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem. The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of the instance.

@@

In general, our approach falls short of the state of the art. 
However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$. Also, there were some instances with high $|\mathcal Q|$ and low $|\mathcal X|$ that were solved by our approach but not the previous solvers.

@@

Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers. Note that ileanCoP and nanoCop were combined for presentation purposes but they also have a significant symmetric difference, e.g. of the 417 problem not solved by our approach there were 68 solved only by ileanCoP and 55 solved only by nanoCoP.

We conclude that our approach already has merits when $|\mathcal Q| = 0$ and $|\mathcal X|$ is small.
We see that encoding idempotence does help to solve some additional benchmarks, but in general does not improve performance.
However, idempotence is a quite powerful concept from a model-theoretic perspective. A more clever encoding of this property could improve results. There are other restrictions on models we could consider, based on structural properties of the formula, e.g. the relation between occurrences of different atoms.
